---
title: "Red Wine Dataset Analysis"
author: "Ola Semsemiah"
date: "1/2/2019"
---

========================================================


```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library("ggplot2")
library("dplyr")
library("gridExtra")
library("ggpubr")

```

```{r echo=FALSE, Load_the_Data,warning=FALSE,message=FALSE}
# Load the Data
Rwine <- read.csv('wineQualityReds.csv')
summary(Rwine)

```




# Univariate Plots Section


```{r echo=FALSE,warning=FALSE,message=FALSE}

# Quality
hist(Rwine$quality,xlab="Quality degree", ylab="Wine", main = 'Wine Quality ',col = 'firebrick')

```


Red wine quality is normally distributed 


```{r echo=FALSE,warning=FALSE,message=FALSE}
#Density
hist(Rwine$density,xlab="Density level", ylab="Wine", main = 'Wine Density ',col = '#DC143C')
```


Red wine density is normally distributed 


```{r echo=FALSE,warning=FALSE,message=FALSE}
# ph level
hist(Rwine$pH,xlab="pH level", ylab="Wine", main = 'Wine pH level ',col = 'indianred')

```


Red wine ph level is normally distributed 


```{r echo=FALSE,warning=FALSE,message=FALSE}
# alcohol
hist(Rwine$alcohol ,xlab="alcohol ", ylab="Wine", main = 'Alcohol  ',col = '#F08080')

```

The distribution of alcohol is right skewed

```{r echo=FALSE,warning=FALSE,message=FALSE}
#fixed.acidity
plot(Rwine$fixed.acidity,  main="fixed.acidity", 
   xlab="wine" , ylab=(expression(fixed~acidity)), pch=6, col=(c("dark blue")))
```


Red wine fixed acidity is right skewed  


```{r echo=FALSE,warning=FALSE,message=FALSE}
#volatile.acidity
plot(Rwine$volatile.acidity,  main="volatile.acidity", 
   xlab="wine" , ylab=(expression(volatile~acidity)), pch=6, col=(c("blue")))

```


Red wine volatile acidity distribution is unclear (bimodal or unimodel, right skewed or normal).


```{r echo=FALSE,warning=FALSE,message=FALSE}
#citric.acid
plot(Rwine$citric.acid,  main="citric.acid", 
   xlab="wine" , ylab=(expression(citric~acidity)), pch=6, col=(c("lightblue")))

```


Red wine citric acid distribution is not normal .


```{r echo=FALSE,warning=FALSE,message=FALSE}
#free.sulfur.dioxide
plot(Rwine$free.sulfur.dioxide,  main="free.sulfur.dioxide", 
   xlab="wine" , ylab=(expression(free~sulfur~dioxide)), pch=9, col=(c("#F0E68C")))


```
```{r echo=FALSE,warning=FALSE,message=FALSE}
#total.sulfur.dioxide
plot(Rwine$total.sulfur.dioxide,  main="total.sulfur.dioxide", 
   xlab="wine" , ylab=(expression(total~sulfur~dioxide)), pch=9, col=(c("#BDB76B")))

```


Both total and free dioxide distribuation are right skewed   


```{r echo=FALSE,warning=FALSE,message=FALSE}
# chlorides
hist(Rwine$chlorides ,xlab="chlorides ", ylab="Wine", main = 'chlorides  ',col = 'darkolivegreen')

```


Red wine chlorides distribution normal .


```{r echo=FALSE,warning=FALSE,message=FALSE}
# residual.sugar 
hist(Rwine$residual.sugar  ,xlab="residual.sugar", ylab="Wine", main = 'residual.sugar ',col = '#808000')


```


The distribution of residual sugar is right skewed


```{r echo=FALSE,warning=FALSE,message=FALSE}
# sulphates
hist(Rwine$sulphates ,xlab="sulphates ", ylab="Wine", main = 'sulphates',col = '#6B8E23')

```


The distribution of sulphates is right skewed 



### What is the structure of your dataset?

Data set rows represnt 1,599 red wines 
Data Set 13 colums represnt red wine features on the chemical properties of the wine (featers are plotted in the previoes section )


### What is/are the main feature(s) of interest in your dataset?

Quality

75% of wine have quality degree that is lower than 6.


### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

Volatile acidity,pH, alcohol


# Bivariate Plots Section


```{r echo=FALSE,warning=FALSE,message=FALSE}
# Alcohol and Quality
ggplot(Rwine,
  aes(y =alcohol , x = quality))+
  geom_jitter(width = 0.25, height = 0, alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE,size=1)+
    ggtitle(' Alcohol vs. Quality')+
    xlab("Quality ") +
  ylab("Alcohol ")

```


Relationship between Alchoal and wine quality


```{r echo=FALSE,warning=FALSE,message=FALSE}
# pH and Quality
ggplot(Rwine,
  aes(y =pH , x = quality))+
  geom_jitter(width = 0.25, height = 0, alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE,size=1)+
    ggtitle(' pH vs.Quality')+
    xlab("Quality") +
  ylab("pH")

```


Relationship between pH and wine quality


```{r echo=FALSE,warning=FALSE,message=FALSE}
#Volatile acidity and Quality
ggplot(Rwine,aes(y =volatile.acidity, x = quality))+
  geom_jitter(width = 0.25, height = 0, alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE,size=1)+
    ggtitle(' Volatile acidity vs.Quality')+
    xlab("Quality") +
  ylab("Volatile acidity")

```


Relationship between Volatile acidity and wine quality


# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?



There is a positive relationship between quality degree and alcohol. 



### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?
There is a negative relationship between quality degree and volatile acidity



### What was the strongest relationship you found?


Quality vs Alcohol 

# Multivariate Plots Section


```{r echo=FALSE,warning=FALSE,message=FALSE}
# Alcohol, Volatile Acidity and  Quality
ggplot(Rwine,
  aes(y =alcohol , x = volatile.acidity, color = as.factor(quality))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE,size=1)+
  
    ggtitle('Alcohol, Volatile Acidity and  Quality')+
    xlab((expression(Volatile~Acidity)))   +
  ylab("Alcohol")
```

Relationship between Alcohol, Volatile Acidity and  Quality


```{r echo=FALSE,warning=FALSE,message=FALSE}
#Alcohol ,pH and by Quality
ggplot(Rwine,
  aes(y =alcohol , x = pH, color = as.factor(quality))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE,size=1)+
  scale_color_brewer(palette = "RdYlBu", name="Quality", direction=-1)+
    ggtitle('Alcohol ,pH and by Quality')+
    xlab('pH')   +
  ylab("Alcohol")

```


Relationship between Alcohol ,pH and by Quality

```{r echo=FALSE,warning=FALSE,message=FALSE}
# Alcohol , density and by Quality
ggplot(Rwine,
  aes(y =alcohol , x = density, color = as.factor(quality))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE,size=1)+
  scale_color_brewer(palette = "RdYlBu", name="Quality", direction=-1)+
    ggtitle('Alcohol , density and by Quality')+
    xlab('density')   +
  ylab("Alcohol")

```

Relationship between Alcohol , density and by Quality



# Multivariate Analysis

1- Density does not play a role in changing the quality with constant Alcohol
2- More alcohol and less pH or less Volatile Acidity constration produce more quality 

------

# Final Plots and Summary

### Plot One

```{r echo=FALSE, Plot_One}
ggplot(Rwine,
  aes(y =alcohol , x = quality))+
  geom_jitter(width = 0.25, height = 0, alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE,size=1)+
    ggtitle(' Alcohol vs. Quality')+
    xlab("Quality ") +
  ylab("Alcohol ")
```


Alcohol percentage has played a big role in determining the quality of Wines


### Plot Two


```{r echo=FALSE,warning=FALSE,message=FALSE}
ggplot(Rwine,
  aes(y =alcohol , x = pH, color = as.factor(quality))) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE,size=1)+
  scale_color_brewer(palette = "RdYlBu", name="Quality", direction=-1)+
    ggtitle('Alcohol ,pH and by Quality')+
    xlab('pH')   +
  ylab("Alcohol")

```

pH affect the quality in a negitive way , More alcohol but less pH produce more quality 


### Plot Three

```{r echo=FALSE,warning=FALSE,message=FALSE, Plot_Three}

hist(Rwine$quality,xlab="Quality degree", ylab="Wine", main = 'Wine Quality ',col = 'firebrick')
```

our dataset contains 'Average' quality wines and there is not too many data in the extreme ranges, thus it is normally distibuted 



------

# Reflection

- In this project we showed that there are many factors that affect the quality of wine, incresing the level 
of some factors might have negative or postive effect to the quality dapanding on the factor type 

- for this project I needed to see some extra resourses for better understanding 

- In genral dataset were enough for predicting the wine quality 

- For future analysis, I would love to have a dataset having a rank for the wine by different tasters so I can explore how humen elemnt can affect the quality 

